---
title: "Netflix EDA"
author: "Nurul Hafizah Zaini"
date: "2025-06-30"
output: html_document
---

## Introduction
### Dataset 

The dataset was obtained from an open-source Kaggle platform. This dataset consists of listings of all the movies and tv shows available on Netflix, along with details such as - cast, directors, ratings, release year, duration, etc.

### Project Context

The project aims to explore trends on type distribution, genres over time, and missing values. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 8, fig.height = 5)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
```

## Load Dataset

```{r, warning=FALSE, echo = FALSE}
df <- read_csv("data/netflix_titles.csv")
```

## Dataset Overview
### Dataset head
```{r, warning=FALSE, echo = FALSE}
head(df)
```

### Dataset summary
```{r, warning=FALSE, echo = FALSE}
summary(df)
```

### Dataset dimension
```{r, warning=FALSE, echo = FALSE}
dim(df)
```

### Columns' names
```{r, warning=FALSE, echo = FALSE}
names(df)
```

### Check for missing Value

```{r, warning=FALSE, echo = FALSE}
colSums(is.na(df))
```

## Exploratory Data Analysis (EDA)
### Movies vs Tv Shows

```{r, warning=FALSE, echo = FALSE}
df %>%
  count(type) %>%
  ggplot(aes(x = type, y = n, fill = type)) +
  geom_col() +
  labs(title = "Distribution by Type", y = "Count")
```

Key insight: The majority of Netflix catalog is Movie type. It makes up about 2/3 of the Netflix titles.  

### Titles Added by Year

```{r, warning=FALSE, echo = FALSE}
df %>%
  mutate(date_added = mdy(date_added),
         year_added = year(date_added)) %>%
  count(year_added) %>%
  ggplot(aes(x = year_added, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Number of Titles Added per Year",
       x = "Year",
       y = "Count")
```

Key insight: The number of new titles increased sharply from 2016 to 2019. A slight reduction in 2020–2021 may reflect COVID-19 impacts on production and acquisitions. 

### Top Genres Per Year

```{r, warning=FALSE, echo = FALSE}
genre_year <- df %>%
  mutate(
    date_added = lubridate::mdy(date_added),
    year_added = lubridate::year(date_added)
  ) %>%
  filter(!is.na(listed_in), !is.na(year_added), year_added >= 2016, year_added <= 2021) %>%
  separate_rows(listed_in, sep = ", ")

top_genres <- genre_year %>%
  group_by(year_added, listed_in) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(year_added) %>%
  slice_max(order_by = count, n = 5) %>%
  ungroup()

ggplot(top_genres, aes(x = reorder(listed_in, count), y = count, fill = listed_in)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ year_added) +
  labs(title = "Top 5 Genres by Year (2016–2021)", x = "Genre", y = "Count")

```

Key insight: Consistently popular genre are International Movies, Dramas, Comedies and International TV Shows. 

## Genres Trend Over The Year
```{r, warning=FALSE, echo = FALSE}
selected_genres <- c("Dramas", "Comedies", "Documentaries", "Action & Adventure")

genre_trend <- top_genres %>%
  filter(listed_in %in% selected_genres)

ggplot(genre_trend, aes(x = year_added, y = count, color = listed_in)) +
  geom_line(linewidth = 1.2) +
  geom_point() +
  labs(title = "Selected Genre Trends Over Time",
       x = "Year", y = "Number of Titles", color = "Genre") +
  theme_minimal()
```

Key insight: 

- Dramas saw a sharp increase between 2016 and 2019, peaking in 2019 with over 550 titles added, before a slight decline in 2020–2021.

- Comedies followed a similar pattern, with steady growth up to 2019 and a noticeable drop in 2021.

- Documentaries rose rapidly between 2016–2017 but plateaued afterward, indicating early investment that later stabilized.

- Action & Adventure remained relatively flat across all years, suggesting consistent but modest investment in this genre.

- The overall decline across genres in 2020–2021 may reflect production slowdowns during the COVID-19 pandemic.